A user-extensible and adaptable parser architecture

### 2008-08-15 doc/sgai-2008.tex
Removed some unused labels cluttering the paper. Minor cleanup.

% $Id$
\documentclass[draft]{svmult}

@@

XXX RENAME ``REJECTION'' TO ``MAIL\_REJECTED'' AND ``IGNORE'' TO
``UNINTERESTING\_LOG\_LINE'' OR SOMETHING ELSE THAT MAKES IT BRUTALLY
OBVIOUS THE PARSER IS SUCCESSFULLY DEALING WITH THE LOG LINES\@.  I SHOULD
PROBABLY RENAME ``SAVE\_BY\_QUEUEID'' TOO\@.

Each action contains the code required to deal with a particular type of
log lines, e.g.\ rejections.  The actions are parser-specific: each parser
author will need to write the actions required from scratch, unless
extending an existing parser.  In the Postfix log parser developed for this
project there are \numberOFactions{} actions and \numberOFrules{} rules,
with a very skewed distribution of rules to actions as shown in
\graphref{Distribution of rules per action}.  The distribution is skewed
for the same reason the architecture is so successful in enabling users to
parse new log lines: most new log lines can be parsed by a combination of a
new rule and an existing action.  It is expected that parsers utilising
this architecture will have a high ratio of rules to actions, hopefully
resulting in simpler rules and clearer distinctions between log lines.
\showgraph{build/graph-action-distribution}{Distribution of rules per
action}{Distribution of rules per action} Unsurprisingly the action with
the highest number of associated rules is \texttt{DELIVERY\_REJECTED}, the action
which handles Postfix rejecting a mail; it is followed by
\texttt{SAVE\_DATA}, the action responsible for handling informative
inputs: these are log lines that supplement the data gathered from other
log lines (such as rejections), but that do not by themselves represent
interesting events.  The third most common action is, perhaps surprisingly,
\texttt{UNINTERESTING}: this action does nothing when executed,
allowing uninteresting log lines to be parsed without causing any effects
(it does not imply that the input is
ungrammatical or unparsed).  The framework warns about every unparsed log
line, to alert the user that they need to alter or extend their ruleset;
the Postfix log parser successfully parses every input line in the
\numberOFlogFILESall{} log files it is currently being tested with.  The
remaining actions have only one or two associated rules: in some cases this
is because there will only ever be one type of log line for that action,
e.g.\ all log lines showing that a remote client has connected are matched
by a single rule and handled by the \texttt{CONNECT} action; in other cases
the action is required to address a deficiency in the log files, or a
complication which arises during parsing.  Taking the \texttt{CONNECT}
action as an example, it creates a new data structure for each client
connecting, and saves the data extracted by the rule (in memory,
subsequently to be entered into the database).  If the data structure
already exists it is treated as a symptom of a bug, and the action issues a
warning containing the full contents of the data structure, plus the log
line which has just been parsed.

@@

\label{complications}

@@

\label{overlapping rules}

@@

\label{example rule}

@@

\label{Rule ordering}
\label{rule ordering for efficiency}
\label{rule efficiency}

>>>>

% $Id$
% XXX REMOVE draft OPTION
\documentclass[draft]{svmult}

@@

Each action contains the code required to deal with a particular type of
log lines, e.g.\ rejections.  The actions are parser-specific: each parser
author will need to write the actions required from scratch, unless
extending an existing parser.  In the Postfix log parser developed for this
project there are \numberOFactions{} actions and \numberOFrules{} rules,
with a very skewed distribution of rules to actions as shown in
\graphref{Distribution of rules per action}.  The distribution is skewed
for the same reason the architecture is so successful in enabling users to
parse new log lines: most new log lines can be parsed by a combination of a
new rule and an existing action.  It is expected that parsers utilising
this architecture will have a high ratio of rules to actions, hopefully
resulting in simpler rules and clearer distinctions between log lines.
\showgraph{build/graph-action-distribution}{Distribution of rules per
action}{Distribution of rules per action} Unsurprisingly the action with
the highest number of associated rules is \texttt{DELIVERY\_REJECTED}, the
action which handles Postfix rejecting a mail; it is followed by
\texttt{SAVE\_DATA}, the action responsible for handling informative
inputs: these are log lines that supplement the data gathered from other
log lines, but that do not by themselves represent interesting events.  The
third most common action is, perhaps surprisingly, \texttt{UNINTERESTING}:
this action does nothing when executed, allowing uninteresting log lines to
be parsed without causing any effects (it does not imply that the input is
ungrammatical or unparsed).  Generally rules specifying the
\texttt{UNINTERESTING} action parse log lines which are not associated with
a specific mail, e.g.\ warnings about configuration files changing.  The
framework warns about every unparsed log line, to alert the user that they
need to alter or extend their ruleset; the Postfix log parser successfully
parses every input line in the \numberOFlogFILESall{} log files it is
currently being tested with.  The remaining actions have only one or two
associated rules: in some cases this is because there will only ever be one
type of log line for that action, e.g.\ all log lines showing that a remote
client has connected are matched by a single rule and handled by the
\texttt{CONNECT} action; in other cases the action is required to address a
deficiency in the log files, or a complication which arises during parsing.
Taking the \texttt{CONNECT} action as an example, it creates a new data
structure for each client connecting, and saves the data extracted by the
rule (in memory, subsequently to be entered into the database).  If the
data structure already exists it is treated as a symptom of a bug, and the
action issues a warning containing the full contents of the data structure,
plus the log line which has just been parsed.

@@

@@

@@

@@

\label{Rule ordering}
